---
title: "Ukraine IDPs and War Fire Events"
author: "Andrea Nasuto"
format: html
editor: visual
---

## War Fire

Here Some Description

## Code

### Load Data

```{r}
# load required package
library(dplyr)
library(tidyr)
library(lubridate)
library(sf)
library(ggplot2)
library(ggmap)
library(tmap)
library(tmaptools)
library(ggspatial)
library(corrr)
library(plotly)
```

### Load Data

```{r}
# read the csv file
war_fires <- read.csv("data/war_fire/ukraine_war_fires.csv")
```

```{r}
head(war_fires)
```

```{r}
# convert date column to date type
war_fires$date <- ymd(war_fires$date)
```

### Select data with Feb - Aug 2022

```{r}
# get the earliest date of the dataset
print(min(war_fires$date))
```

```{r}
# define start and end dates
start_date <- ymd("2022-02-24")
end_date <- ymd("2022-08-31")

# filter activities within the date range
war_fires_filtered <- war_fires %>%
  filter(date >= start_date & date <= end_date)

# display the first few rows of the filtered data
head(war_fires_filtered)
```

```{r}
# read shapefiles
oblasts <- st_read("data/shapefiles/oblasts/oblasts.shp") 
raions <- st_read("data/shapefiles/raions/raions.shp")

#oblasts <- st_read("data/gadm41_UKR_shp/gadm41_UKR_1.shp")[-1,] # skip first row
#raions <- st_read("data/gadm41_UKR_shp/gadm41_UKR_2.shp")[-1,] # skip first row
```

```{r}
# create weeks df
war_fires_filtered <- war_fires_filtered %>% mutate(week = floor_date(date, "week", week_start = 1))
weeks <- unique(war_fires_filtered$week)
weeks <- unique(war_fires_filtered$week) %>% sort()
weeks <- data.frame(week = weeks, week_number = seq_along(weeks))

# create dataset with all weeks and oblasts and raions name
all_oblast_weeks <- expand.grid(
  oblast_name = unique(oblasts$ADM1_EN), #ADM1_EN for correct raions
  week = weeks$week
)

# get the week number
all_oblast_weeks <- merge(all_oblast_weeks, weeks %>% select(week, week_number), by = "week")

# raion
all_raion_weeks <- expand.grid(
  raion_name = unique(raions$ADM2_EN), #ADM2_EN for correct raions
  week = weeks$week
)
# get the week number
all_raion_weeks <- merge(all_raion_weeks, weeks %>% select(week, week_number), by = "week")
```

```{r}
# create spatial dataset from war_fires_filtered
war_fires_filtered_geom <- st_as_sf(war_fires_filtered, coords = c("LONGITUDE", "LATITUDE"), crs = 4326) # set crs to WGS84

# ensure geometry is valid
war_fires_filtered_geom = sf::st_make_valid(war_fires_filtered_geom)

war_fires_filtered_geom <- st_transform(war_fires_filtered_geom, crs = 4326)  # Use appropriate CRS
```

```{r eval=FALSE, include=FALSE}
# set tmap to interactive mode
tmap_mode("view")

# Create the interactive map
war_fire_map <- tm_shape(war_fires_filtered_geom) +
  tm_basemap(server = "OpenStreetMap") +
  tm_dots() +
  tm_layout(title = "War-Fire Events")

war_fire_map <- tm_shape(war_fires_filtered_geom) +
  tm_basemap(server = "OpenStreetMap") +
  tm_dots() +
  tm_layout(title = "War-Fire Events")

# Create the tmap element
raions_GADM_map <- tm_shape(raions) +
  tm_borders() +  # Only show boundaries
  tm_fill(alpha = 0)  # Make the fill transparent

#war_fire_map + raions_GADM_map
```

### Aggregate war-fire events at raion and oblast level

```{r warning=FALSE, message= FALSE}
# rename war_fires_filtered_geom
wf <- war_fires_filtered_geom

# ensure both shapefiles and events are in the same CRS
oblasts <- st_transform(oblasts, st_crs(wf))
raions <- st_transform(raions, st_crs(oblasts))

# spatial join to add oblast and raion info to the events
wf_oblasts <- st_join(oblasts, wf, left = TRUE) # deselect date with correct raions shp
wf_raions <- st_join(raions, wf, left = TRUE) # deselect date with correct raions shp

# rename area name variables
wf_oblasts <- wf_oblasts %>% rename(oblast_name = ADM1_EN) #ADM2_EN for correct raions
wf_raions <- wf_raions %>% rename(raion_name = ADM2_EN) #ADM2_EN for correct raions

# aggregate by oblast and week
oblast_weekly <- wf_oblasts %>%
  group_by(oblast_name, week) %>%
  summarise(count = n(), .groups = 'drop')

# aggregate by raion and week
raion_weekly <- wf_raions %>%
  group_by(raion_name, week) %>%
  summarise(count = n(), .groups = 'drop')
```

```{r}
# include all weeks and oblasts in the aggregated df per oblast per week
oblast_weekly <- left_join(all_oblast_weeks, oblast_weekly, by = c('oblast_name', 'week'))

# when oblasts have NA replace with 0 since it means that there are no events
oblast_weekly$count[is.na(oblast_weekly$count)] <- 0

# attach geometries since some oblasts have empty geometry
oblast_weekly <- left_join(oblast_weekly %>% select(-geometry), oblasts %>% select(ADM1_EN, geometry), by = c("oblast_name" = "ADM1_EN"))

# include all weeks and raions in the aggregated df per raion per week
raion_weekly <- left_join(all_raion_weeks, raion_weekly, by = c('raion_name', 'week'))

# when oblasts have NA replace with 0 since it means that there are no events
raion_weekly$count[is.na(raion_weekly$count)] <- 0

# attach geometries since some oblasts have empty geometry
raion_weekly <- left_join(raion_weekly %>% select(-geometry), raions %>% select(ADM2_EN, geometry), by = c("raion_name" = "ADM2_EN"))
```

```{r}
# convert the 'week' column to character format for raion_weekly and oblast_weekly
raion_weekly$week <- as.character(raion_weekly$week)
oblast_weekly$week <- as.character(oblast_weekly$week)

# convert the 'week' column to Date format using lubridate for raion_weekly and oblast_weekly
raion_weekly$week <- ymd(raion_weekly$week)
oblast_weekly$week <- ymd(oblast_weekly$week)
```

#### Standardization

```{r}
# min-max standardization
raion_weekly <-  raion_weekly %>%
  mutate(intensity = (count - min(count, na.rm = TRUE)) / (max(count, na.rm = TRUE) - min(count, na.rm = TRUE)) * 100)
```

```{r}
# min-max standardization
oblast_weekly <-  oblast_weekly %>%
  mutate(intensity = (count - min(count, na.rm = TRUE)) / (max(count, na.rm = TRUE) - min(count, na.rm = TRUE)) * 100)
```

```{r}
# z-score standardization and then min-max to fit the 0-100 scale

# z-score standardization
raion_weekly <- raion_weekly %>%
  mutate(z_score = (count - mean(count, na.rm = TRUE)) / sd(count, na.rm = TRUE))

# min-max normalization to scale z-scores between 0 and 100
raion_weekly <- raion_weekly %>%
  mutate(intensity = ((z_score - min(z_score, na.rm = TRUE)) / 
                      (max(z_score, na.rm = TRUE) - min(z_score, na.rm = TRUE))) * 100)
```

#### Export datasets

```{r}
write.csv(raion_weekly %>% select(-geometry), "output/data/war_fire_raion.csv", row.names = FALSE) # export war fire events at raion level
```

```{r}
write.csv(oblast_weekly %>% select(-geometry), "output/data/war_fire_oblast.csv", row.names = FALSE) # export war fire events at oblast level
```

## Visualizations

```{r}
# Create the map for raion_month
raion_map <- tm_shape(raion_month) +
  tm_basemap(server = "OpenStreetMap") +
  tm_polygons("count", palette = "Blues", title = paste0("Raion-level Monthly Count - ", '0', month))

# Create the map for oblast_month
oblast_map <- tm_shape(oblast_month) +
  tm_basemap(server = "OpenStreetMap") +
  tm_polygons("count", palette = "Reds", title = paste0("Oblast-level Monthly Count - ", '0', month))

# Combine both maps
combined_map <- oblast_map + raion_map

# set tmap to interactive mode
tmap_mode("view")
tmap::tmap_leaflet(combined_map)
```

#### Map Animation at Raions and Oblast level for each week

```{r}
# make sure week is a date
raion_weekly$week <- as.Date(raion_weekly$week)

# check no NA in week
raion_weekly_animation <- raion_weekly %>% filter(!is.na(week))

# set dataset as sf
raion_weekly_animation <- st_as_sf(raion_weekly_animation, sf_column_name = "geometry", crs= 4326)

# arrange from the latest to the earliest
raion_weekly_animation <- raion_weekly_animation %>% arrange(week)

# create breaks
breaks <- classInt::classIntervals(raion_weekly_animation$intensity, n = 5, style = "jenks")$brks

# Function to create individual maps
create_map <- function(time) {
  tm_shape(raion_weekly_animation %>%
             filter(week == time)) +
    tm_basemap("OpenStreetMap") +
    tm_polygons("intensity", palette = "Blues", title = "Intensity (0-100)", 
                breaks = breaks) +
    tm_layout(title = paste("Week:", time))
}

# List of time points
time_points <- unique(raion_weekly_animation$week)

# Create the animation
tmap_animation(
  tm = lapply(time_points, create_map),
  filename = "output/figures/old_raion_intensity.gif",
  width = 800,
  height = 600,
  delay = 100
)
```

### Interactive Map at Raion Level

```{r}
time = '2022-07-11'
tm_shape(raion_weekly_animation %>%
             filter(week == time)) + # select specific week by looking at time_points
    tm_polygons("intensity", palette = "Blues", title = "Intensity (0-100)", 
                breaks = breaks) +
    tm_layout(title = paste("Week:", time))

tmap_mode("view")
```

## Internal Displacement and War-Fire Events

### IDPs data

```{r}
# load the internal displacement data grouped at raion level (new boundaries)
idp_raion <- read.csv('data/displacement_data/updated_raions.csv')
```

```{r}
# rename raion variable
idp_raion <- idp_raion %>% rename(raion_name = ADM2_EN,
                                  oblast_name = ADM1_EN)

# calculate the week by date

```

### Relationship between IDPs data and war-fire intensity

"net_refugee_outflow" is the net number of people leaving Ukraine on a particular day after the invasion\
\
"net_pop_change" is the

"proportional_net_change" is the

```{r}
# merge the datasets by raion_name and week_number
merged_data <- idp_raion %>%
  inner_join(raion_weekly, by = c("raion_name", "week_number")) %>%
  select(raion_name, week, week_number, counts, intensity, net_refugee_outflow, net_pop_change, proportional_net_change)
```

```{r}
# 1) calculate the correlation matrix

# select relevant columns for correlation
cor_data <- merged_data %>%
  select(net_refugee_outflow, net_pop_change, proportional_net_change, counts, intensity)

# calculate the correlation matrix
cor_matrix <- cor(cor_data, use = "complete.obs")

# display the correlation matrix
print(cor_matrix)

# visualize the correlation matrix
cor_plot <- cor_matrix %>%
  as_cordf() %>%
  network_plot(min_cor = 0.2)
```

```{r}
# 2) display the correlation between war-fire events intensity and population net change

# convert raion_weekly to spatial data
raion_sf <- st_as_sf(raion_weekly, coords = c("longitude", "latitude"), crs = 4326)

# merge spatial data with idp_raion data
map_data <- raion_sf %>%
  inner_join(idp_raion, by = "raion_name")

# plot the map with interaction
tmap_mode("view")
tm_shape(map_data) +
  tm_polygons("net_pop_change", title = "Net Population Change") +
  tm_facets(by = "intensity", free.coords = FALSE)
```

```{r}
# 3) produce a bivariate map displaying intensity and net_pop_change

# create a new variable for bivariate mapping
map_data <- map_data %>%
  mutate(bivariate = cut(net_pop_change, breaks = 5, labels = FALSE) +
           cut(intensity, breaks = 5, labels = FALSE) * 10)

# plot the bivariate map
tmap_mode("view")
tm_shape(map_data) +
  tm_polygons("bivariate", title = "Bivariate Map") +
  tm_facets(by = "bivariate", free.coords = FALSE)
```

```{r}
# 4) produce a scatterplot between intensity and net_pop_change

# plot the scatterplot
scatter_plot <- ggplot(merged_data, aes(x = intensity, y = net_pop_change)) +
  geom_point() +
  labs(title = "Scatterplot between War-Fire Intensity and Net Population Change",
       x = "Intensity",
       y = "Net Population Change")

# make the scatterplot interactive
interactive_scatter <- ggplotly(scatter_plot)

# display the interactive scatterplot
interactive_scatter
```

### Code Dump

```{r}
# remove duplicate - !this is not required with with st_join centroids
idp_raion_origin_unique <- idp_raion_origin %>% distinct(ADM2_EN, ADM2_EN, month, day, .keep_all = TRUE)

idp_raion_destination_unique <- idp_raion_destination %>% distinct(ADM2_EN, ADM2_EN, month, day, .keep_all = TRUE)

# set up dataset as sf
idp_raion_origin <- st_as_sf(idp_raion_origin_unique, sf_column_name = "geometry", crs= 4326)

idp_raion_destination <- st_as_sf(idp_raion_destination_unique, sf_column_name = "geometry", crs= 4326)
```

```{r}
# spatial joint idp raion at origin with new raion shapefiles
idp_raion_new_origin <- st_join(raions %>% select(1,5, geometry), idp_raion_origin, 
                         #largest = TRUE
)

idp_raion_new_destination <- st_join(raions %>% select(1,5, geometry), idp_raion_destination)
#                         #largest = TRUE)
```

```{r}
# create a unique dataset with ADM2_EN and ADM2_EN from idp_raion_new_destination
unique_pairs <- idp_raion_new_destination %>%
  st_drop_geometry() %>%
  select(ADM2_EN, ADM2_EN) %>%
  distinct() %>%
  na.omit()

# rename ADM2_EN to raion_destination_name
unique_pairs <- unique_pairs %>%
  rename(raion_name_destination = ADM2_EN)
```

```{r}
# match name of the old raion with the new raion for the destination
idp_raion_new <- idp_raion_new_origin %>% left_join(unique_pairs, by = "ADM2_EN")
```

```{r}
summarized_values <- idp_raion_new %>% st_drop_geometry() %>%
  group_by(raion_name_origin, raion_name_destination, month, day) %>%
  summarise(
    total_baseline_raion_population = sum(total_baseline_raion_population, na.rm = TRUE),
    total_count = sum(total_count, na.rm = TRUE),
    net_pop_change = sum(net_pop_change, na.rm = TRUE)
  )
```
